\makeatletter
\ifthenelse{\pdf@strcmp{\languagename}{english}=0}
{\addcontentsline{toc}{chapter}{Acknowledgments}}
{\addcontentsline{toc}{chapter}{Danksagungen}}
\makeatother
\thispagestyle{empty}

\vspace*{20mm}

\begin{center}
\makeatletter
\ifthenelse{\pdf@strcmp{\languagename}{english}=0}
{\usekomafont{section} Acknowledgments}
{\usekomafont{section} Danksagungen}
\makeatother
\end{center}

\vspace{10mm}

I would like to say thanks to all who supported me in the last few months, to all who endured my mood, and to all, who encouraged me to stay motivated. Thank you!

I would like to take this opportunity to thank my advisor Peter Schüffler for giving me the opportunity to write my thesis at his chair, allowing me to explore my topic to the fullest, and taking time for numerous meetings and detailed discussions. Thank you and hope we will meet in person someday.

Special thanks also go to all Munich AstraZeneca members for providing amazing resources and constant support along the way. Many thanks to my advisors, Philipp Wortmann and Ansh Kapil for trusting me to work on a remarkable topic. I have learned a lot, thanks to you and your guidance. Particular thanks go to Armin Meier who always found time to discuss frustrating details and provided elegant solutions every time. And thanks to all other sync meetings and lunch break discussions with Guillaume Potdevin, Abdullah Hayran, Gordan Prastalo, Bianca Mocanu, and Srividhya Sathya Narayanan. 

Finally, I would like to say thanks for all the non-scientific support. To my family, especially my parents and my grandmother, who repeatedly reminded me not to work too hard.
Ukrainian Armed Forces and all defenders and volunteers for protection and the chance to do science. To all Ukrainians for sharing these difficult times with me and reminding me that there should still be room for normal even in war times.  And thanks to my wonderful friends Lena Shevtsova and Laczik sisters, Dalma and Dori for being there for me.

Thank you.

\cleardoublepage{}
